---
title: "General Social Survey"
author: "Sandra Iovno   "
date: "Summer 2024"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```
 
#### Introduction 

 This data came from GSS.data. The background death penalty (CAPPUN) independent of gun ownership (OWNGUN). 
 
 $H_0$: CAPPUN is independent of OWNGUN   
$H_A$ : That they are independent 

 
#  Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  
First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?



## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
 
  CAPPUN is a variable,  
  OWNGUN is a variable,
  so I use the helper file    
  CAT ~ CAT  
    
    
##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

```{r}
barchartGC(~CAPPUN + OWNGUN,data=GSSdata)
barchartGC(~CAPPUN + OWNGUN,data=GSSdata, type="percent")

```

  This graph shows that the people who are in favor of capital punishment is about 50/50 in owning a gun. 
    The people who appose capital punishment are less likely to own a gun which makes it dependent. 

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

```{r}
table1 <- xtabs(~CAPPUN + OWNGUN, data=GSSdata)
rowPerc(table1)
colPerc(table1)
```
  This table shows that people who are in favor of the capital punishment is about 76% own a gun.  
    Where poeple who are ppose to it is about 42% who own a gun.  


<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

 Chi-squared shows P-Value of 0.02 which is less than 0.05 so we reject this hypothesis CAPPUN is independent of OWNGUN.  
 
 The fisher's Exact test P- Value is 0.01 but we round up to 0.02 which is still less than 0.05.  
 This in means that fishers data rejects the null hypothesis as well as with the CHI-Square test . The odds of the ratio is 2.271 : 1 which means that gun owners are 2.3 times more likely to favor capital punishment than those who don't.   
 
 
### Inconclusion 

We tried to answer the question that was giving to us " Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)? We when doing the graphical charts we discovered that this graph shows that the people who are in favor of capital punishment is about 50/50 in owning a gun. 
  The people who appose capital punishment are less likely to own a gun which makes it dependent.
  When doing the chi-sqaure the P-Value was 0.02 which is less than 0.05. This indicates that we reject the hypothesis. This means that they are that they are independent. 

#  Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?


#### Introduction 
We got this file from GSS.data file.  
$H_0$ : There is no difference of years EDU who did not VOTE12
$H_A$ : There is a difference of years EDUC who did VOTE12

## Methods
EDUC is Num  
VOTE12 is a Factor   
So I used helper file  
NUM ~ CAT 


<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

## Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results
<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
plot(EDUC~VOTE12,data=GSSdata)
```

  This boxchart shows that the people who votes in 2012 tend to have higher EDUC than those who did not VOTE12.       
  
  
#### Numerical Descriptive Results
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
favstats(EDUC~VOTE12,data=GSSdata)
```
This Numerical Results shows that in 2012 those who did vote have almost 16 years in education.  
Those who did not vote in 2012 have a little over 13 years of education.  

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
t.test(EDUC~VOTE12,data=GSSdata)
``` 
this T tests shows that this means those who voted in 2012 have higher education than those who did not. The P-Value is 1.248e-07. When solving this equation it would be .0000001 so we reject the null hypothesis there is no difference of years EDU who did not VOTE12. 


### Conclusion  
In this question we tried to figure out how many people voted in 2012 have a higher education.  
When doing the  boxchart shows that the people who voted in 2012 tend to have higher EDUC than those who did not VOTE12. Then when doing the Numerical graph it shows the amount of people who voted in 2012 have 16 years of education which is more than those who didn't vote in 2012. T tests shows that this means those who voted in 2012 have higher education than those who did not. The P-Value is 1.248e-07. When solving this equation it would be .0000001 so we reject the null hypothesis there is no difference of years EDU who did not VOTE12. In conclusion there is a difference in those who did vote 2012 than those who didn't. 

#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->
 
Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

### Introduction  

$H_0$ Wordsum is not independent of level schooling DEGREE    
$H_A$ They are independent   

## Methods
WORDSUM : NUM   
DEGREE : Factor   
helper file I used    
NUM ~ CAT 


<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

## Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
plot(WORDSUM~DEGREE,data=GSSdata)
```
  
This graph shows the Junior College and Bachelor Degree have about the same amount of Wordsum test,  
but graduates have a little more than Junior College or Bachelor Degree.      


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=GSSdata)
```

The numerical description results that graduates have the highest mean and it is less than high school because it is the lowest  

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSSdata)
summary.aov(model1)
```
```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 8.7, 2, 2))
plot(TukeyHSD(model1),las=2)
```

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


 This graph shows that the P-Value is .0000002, which is less than 0.05. This means that we reject the null hypothesis Wordsum  is not independent of level schooling DEGREE. They are in fact independent of one another.    
    
Tukey conclusion states there is no difference high school and less than high school.   
There is a difference between junior College and High school      
no difference between high school and graduates     
no difference between junior college and graduates     
no difference between bachelors an graduates.   
    
#  Conclusion
<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

We tried to answer whether a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE) or not independent. what we came to conclude is that the P-value was less than 0.05, so we rejected the null hypothesis that they are not independent to each other. They are in fact dependent to each other. 
